---
title: "Lab4"
author: "Stella (Yuxin) Wan"
date: "2019/10/10"
output: html_document
---
```{r}
library(XML)
library(rvest)
library(tidyverse)
library(data.table)
```

## Scrape the Location Counts
### 1.1 Make a table of brands and related url
```{r}
coffeeBrands<-data.table(brands=c( 'Starbucks',
                                   'Dunkin Donuts',
                                   'Peets',
                                   'Tim Hortons',
                                   'Panera Bread',
                                   'Caribou Coffee',
                                   'Au Bon Pain',
                                   'The coffee Bean & Tea Leaf',
                                   "McDonald's"),
url= c("https://www.menuism.com/restaurant-locations/starbucks-coffee-39564",
        "https://www.menuism.com/restaurant-locations/dunkin-donuts-181624",
        "https://www.menuism.com/restaurant-locations/peets-coffee-tea-84051",
        "https://www.menuism.com/restaurant-locations/tim-hortons-190025",
        "https://www.menuism.com/restaurant-locations/panera-bread-4258",
        "https://www.menuism.com/restaurant-locations/caribou-coffee-164861",
        "https://www.menuism.com/restaurant-locations/au-bon-pain-69342",
        "https://www.menuism.com/restaurant-locations/the-coffee-bean-tea-leaf-165988",
        "https://www.menuism.com/restaurant-locations/mcdonalds-21019"
))
```

### 1.2 Print location text given url
```{r}
LocationCounts<- function(url){
    link<- read_html(url)
    webpage<- html_nodes(link,css = "section div div div ul li a")
    data<- html_text(webpage)
    return(data)
}
```

### 1.3 Covert to table with state,abb and number of locations given data

#stateabb
```{r}
stateabb <- function(fullStateName){
  return (state.abb[which(state.name == fullStateName)])
}
```

```{r}
statesAndNumberofLocation<- function(data){
  #Extract US states regular expression
  regexStates <- "(Alabama|Alaska|Arizona|Arkansas|California|Colorado|Connecticut|Delaware|Florida|Georgia|Hawaii|Idaho|Illinois|Indiana|Iowa|Kansas|Kentucky|Louisiana|Maine|Maryland|Massachusetts|Michigan|Minnesota|Mississippi|Missouri|Montana|Nebraska|Nevada|New\ Hampshire|New\ Jersey|New\ Mexico|New\ York|North\ Carolina|North\ Dakota|Ohio|Oklahoma|Oregon|Pennsylvania|Rhode\ Island|South\ Carolina|South\ Dakota|Tennessee|Texas|Utah|Vermont|Virginia|Washington|West\ Virginia|Wisconsin|Wyoming)"
  #Extract numbers of locations regular expression
  regexNumbers<- '[0-9]+'
  result <- data.table(state = character(), abb=character(), number = integer())
  #loop every line to ???
  for (i in 1:length(data)){
    #see if it matches a US states
    isInUS<- grepl(regexStates,data[i])
    if (isInUS == TRUE){
      #get state name using regex
      stateMatch <- regexpr(regexStates,data[i])
      fullStateName <- regmatches(data[i],stateMatch)
      #convert to abb
      stateAbb <- stateabb(fullStateName)
      #get number using regex
      numberMatch <- regexpr(regexNumbers,data[i])
      number <- as.integer(regmatches(data[i],numberMatch))
      # add to the table
      result <- rbind(result,data.table(state = fullStateName, abb = stateAbb, number = number))
    }
  }
  return (result)
}
```
    

#For each brand, get number of locations and make a bigg table
```{r}
for (i in 1:nrow(coffeeBrands)){
  rawdata <- LocationCounts(coffeeBrands[i, url])
  table <- statesAndNumberofLocation(rawdata)
  colnames(table)[3] <- coffeeBrands[i, brands]
  if (exists("bigTable")){
    bigTable <- merge(x= bigTable, y= table, by= c("state","abb"), all = TRUE)
  } else{
    bigTable <- table
  }
}
bigTable
```

## Supplemental Data
### 2.1 Get states and population (question 4)
```{r}
statesPopulationurl<-
"https://simple.wikipedia.org/wiki/List_of_U.S._states_by_population"
statesPopulationlink <- read_html(statesPopulationurl)
statesPopulationhtml <- html_nodes(statesPopulationlink, css="table")
statesPoplulationData <- html_table(statesPopulationhtml)[[1]][,3:4]
#Rename the columns
colnames(statesPoplulationData)[1:2] <- c('state','population')
statesPoplulationDT<- data.table(statesPoplulationData)
#Merge state and population table with previous big table
hugeTable<- merge(x= bigTable, y= statesPoplulationDT, all.x = TRUE, by ='state') 
hugeTable
```

### 2.2 Get market cap data (question 5)
```{r}
M<- 10^6
B<- 10^9
hugeTable$Starbucks_mc<- 103.792*B   
hugeTable$Donuts_mc<- 6.398*B
hugeTable$Peets_mc<- 507.07*M
hugeTable$Hortons_mc<- 11.39*B
hugeTable$Bread_mc<- 7.154*B
hugeTable$Coffee_mc<- 325.56*M
hugeTable$Pain_mc<- 100*M
hugeTable$Leaf_mc<- 350*M
hugeTable$McDonald_mc<- 160.347*B
hugeTable
```

### 2.3 Create a region variable (question 6)
```{r}
#Create table with state and region
regionTable <- data.table(state = state.name, region = state.region)
#merge it to the bigTable
hugeTable <- merge(x= hugeTable, y= regionTable, by= c("state"), all.x = TRUE)
hugeTable
```

## Analyze
### Coffee brands of top number of stores in states
3.1 Are some of these chains more prevalent in certain states than others? Possibly depsite having less stores overall? Same questions for regions instead of states. 
```{r}
for (i in 3:11){
  brands <- colnames(hugeTable)[i]
  stateAndBrandNumber <- hugeTable %>% 
    select(1,brands,22)
  #print(stateAndBrandNumber)
  stateAndBrandNumber[order(desc(stateAndBrandNumber[[brands]]))][1:10] %>% 
  print()
}
    ```
 - Starbucks is more prevalent in California than other brands with most stores, and is the most prevalent brand in the west 
 - Dunkin Donuts is more prevalent in Massachusetts than other brands, and is the most prevalent brand in northeast with most stores 
 - Peets is more prevalent in California than in other states although  it does not have many stores  
 - Tim Hortons is more prevalent in Michigan than other brands although  it does not have many stores 
 - Panera Bread is more prevalent in Florida than other brands 
 - Calibou Coffee is more prevalent in Minnesota than other brands 
 - The coffee Bean & Tea Leaf is more prevalent in California than in other states although it does not have many stores 
 - McDonald's is the most prevalent brand in the south and north central 

3.2 Do the market cap data match what you’d expect based on the number and locations of the stores? Why or why not?
```{r}
coffeeMC <- data.table(
  brands=c(
    'Starbucks',
    'Dunkin Donuts',
    'Peets',
    'Tim Hortons',
    'Panera Bread',
    'Caribou Coffee',
    'Au Bon Pain',
    'The coffee Bean & Tea Leaf',
    "McDonald's"),
  MC=c(103.792*B,6.398*B,507.07*M,11.39*B,7.154*B,325.56*M,100*M,350*M,160.347*B))
#Make a table of total location by state
TotalLocationByState<- hugeTable[,3:11][, lapply(.SD, sum, na.rm=T),] %>% 
  gather(brands, totalNumber)
#Merge coffeeMC with TotalLocationByState
NewTable<- merge(x=coffeeMC, y= TotalLocationByState, by='brands', all=TRUE)
arrange(NewTable,desc(MC))
```

 - No. McDonald's,Starbucks and Tim Hortons are the coffee brands with top 3 market value. So I expected them to have most locations as well. However, the number of locations of Tim Hortons is actually not big.

```{r}
write.csv(hugeTable,"C:/Users/Stella/Desktop/temp/hugeTable.csv")
```

